Board Game Recommendation

### 2015-03-02 paper.Rnw
couple more comments

Note for editing: I know this model is a mess at the moment, just getting the general ideas down on paper. 
\hh{this is not nearly as messy as you make it out to be, but you need to make sure to define every little piece of notation as you, because otherwise your reader will be lost.}
\subsubsection{Global Effects}
The motivations for using the formulas that follow can be found in Scalable Collaborative Filtering with Jointly Derived Neighborhood Interpolation Weights (make this a bib reference). The general idea is to remove the global mean from all ratings to get a residual for each observation. We'll group these residuals by game, and then find an effect for that game. Subtracting this effect from the residual, we'll take those residuals and repeat the same for each user.

As we proceed $y_{u,i}$ will denote the rating by user $u$ for game $i$. Then, we have $$y_{u,i} = \mu_{global} + \mu_{i} + \mu_{u} + error$$

However, we'll approach this in steps, as follows:

@@

\begin{center}
\begin{align}
\mu_{global} = \frac{\sum_{u = 1}^{N_{u}}\sum_{i = 1}^{N_{games}}}{1}
\end{align}
\end{center}

\subsubsection{Classifier Aggrigation}
Independent of the global effects, we need to give each user a score in each classification category. We will use 4 types of categorical variables, which will yield (FIND OUT HOW MANY CLASSIFIER VARIABLES). To do this, we can either use the ratings themselves, or the residuals from the global effects process. Let $c_{u,j}$ be the categorical rating for user $u$ in category $j$. Additionally, for each game $i$, there exists a set of classifiers $C_i$ that describe the aspects of the game in things such as mechanics, game family, categories, and subdomains.  Then for each user, we take all of the categorical values for each user and normalize them, so they should all have mean 0 and standard deviation of 1 for each user. Then, within each category, the average rating is calculated. Now, if a user hasn't rated anything in this category we can assign them a score of 0 in that category (what should amount to an average rating). These serve dual purposes: Firstly, we can assess what sorts of games a user rates higher than average most of the time, and game they rate lower than average most of the time. This will allow the algorithm to not only recommend specific games that have a high predicted rating, but will also allow for the user to find categories or mechanics that they enjoy as well. The other purpose is that we now have a larger dataset to compare people than we might have had before, as we are guarantedd to have (FIND OUT HOW MANY CLASSIFIER VARIABLES) for every user.

@@

This algorithm is implemented in R using shiny as a vehhicle for a simple, easy to use user interface. Behind the scenes, the data is stored in a data base with six tables: A player key table, which relates the players information to a unique player userid; A game key table, which relates game information to a unique game id; A ratings table which stores the ratings (a score from 1-10) along with the userid of the person who rated it and the game id which the rating is for; the classifiers table, relates the game id to characteristics about the game that will be useful in the classifier aggrigation and k-nearest-neighbors section; the predicted ratings table, which will store predicted ratigs if ratings have already been fit for a particular user; and the aggregated classifier ratings for each user. This data storage will allow for ease of implementation: we only really need the ratings table and the classifier table to do all the work, and the key tables are really only needed to display the inforation to the user in a format which they will understand.

@@

We first apply the global effects strategy as descibed in the modling section. The optimization of the $\alpha$ parameters will have occred sometime before this and the algoritm will use use those stored values. A weeky or monthly automatic updaing procedure is used to keep these values up to date as more data comes into the system. From the global effects, we then aggregate those based on the classifier aggrigation procedure as described above. These are then stored in the database for future use by the system. After that, before we move onto the k-nearest neighbors implementation, we must first find an optimal value for $k$ using the 10-fold cross validation as described above.

>>>>

Note for editing: I know this model is a mess at the moment, just getting the general ideas down on paper. 
\hh{this is not nearly as messy as you make it out to be, but you need to make sure to define every little piece of notation as you, because otherwise your reader will be lost.}
\subsubsection{Global Effects}
 The general idea is to remove the global mean from all ratings to get a residual for each observation. We'll group these residuals by game, and then find an effect for that game. Subtracting this effect from the residual, we'll take those residuals and repeat the same for each user. The motivations for using the formulas that follow can be found in Scalable Collaborative Filtering with Jointly Derived Neighborhood Interpolation Weights (make this a bib reference, \hh{you still need to paraphrase the motivation here.}).

As we proceed $y_{u,i}$ will denote the rating by user $u$ for game $i$ \hh{what scale is the rating on?}. Then, we have $$y_{u,i} = \mu_{global} + \mu_{i} + \mu_{u} + error$$

However, we'll approach this in steps, as follows: \hh{what is the benefit of using a hierarchical approach?}

@@

\begin{center}
\begin{align}
\mu_{global} = \frac{\sum_{u = 1}^{N_{u}}\sum_{i = 1}^{N_{games}}}{1}
\end{align}
\end{center}
\hh{something went wrong in the formula - what are you summing over?}

\hh{This is a linear approach to estimate ratings -- one of the problems is, that rating is on a finite scale, i.e. it is possible that predictions go outside this interval - what do you do with those predictions?}

\subsubsection{Classifier Aggregation}
Independent of the global effects, we need to give each user a score in each classification category. We will use 4 types of categorical variables, which will yield (FIND OUT HOW MANY CLASSIFIER VARIABLES). To do this, we can either use the ratings themselves, or the residuals from the global effects process. Let $c_{u,j}$ be the categorical rating for user $u$ in category $j$. Additionally, for each game $i$, there exists a set of classifiers $C_i$ that describe the aspects of the game in things such as mechanics, game family, categories, and subdomains.  Then for each user, we take all of the categorical values for each user and normalize them, so they should all have mean 0 and standard deviation of 1 for each user. Then, within each category, the average rating is calculated. Now, if a user hasn't rated anything in this category we can assign them a score of 0 in that category (what should amount to an average rating). These serve dual purposes: Firstly, we can assess what sorts of games a user rates higher than average most of the time, and game they rate lower than average most of the time. This will allow the algorithm to not only recommend specific games that have a high predicted rating, but will also allow for the user to find categories or mechanics that they enjoy as well. The other purpose is that we now have a larger dataset to compare people than we might have had before, as we are guarantedd to have (FIND OUT HOW MANY CLASSIFIER VARIABLES) for every user.

@@

This algorithm is implemented in R using shiny as a vehhicle for a simple, easy to use user interface. Behind the scenes, the data is stored in a data base with six tables: A player key table, which relates the players information to a unique player userid; A game key table, which relates game information to a unique game id; A ratings table which stores the ratings (a score from 1-10) along with the userid of the person who rated it and the game id which the rating is for; the classifiers table, relates the game id to characteristics about the game that will be useful in the classifier aggregation and k-nearest-neighbors section; the predicted ratings table, which will store predicted ratigs if ratings have already been fit for a particular user; and the aggregated classifier ratings for each user. This data storage will allow for ease of implementation: we only really need the ratings table and the classifier table to do all the work, and the key tables are really only needed to display the inforation to the user in a format which they will understand.

@@

We first apply the global effects strategy as descibed in the modling section. The optimization of the $\alpha$ parameters will have occred sometime before this and the algoritm will use use those stored values. A weeky or monthly automatic updaing procedure is used to keep these values up to date as more data comes into the system. From the global effects, we then aggregate those based on the classifier aggregation procedure as described above. These are then stored in the database for future use by the system. After that, before we move onto the k-nearest neighbors implementation, we must first find an optimal value for $k$ using the 10-fold cross validation as described above.
